---
title: 'Midterm: Biostat 626'
author: "Dhajanae Sylvertooth"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Loading needed data and libraries
```{r, message=FALSE}
library(dplyr)
library(corrr) ## for correlation analysis
library(ggcorrplot) 
library(FactoMineR) ## used to perform PCA
library(utils) ## for reading in the data
library(nnet) ## multinomial logistic regression
library(glmnet)
library(ggplot2)

### set working directory here (data must be downloaded in said directory)
setwd("/Users/dj/Documents/Classes/Biostat 626/Midterm/")

training_data <- read.table("training_data.txt", header = TRUE)
testing_data <- read.table("test_data.txt", header = TRUE)

### Data for task 1
training_data.t1 <- training_data %>%
  mutate(status = ifelse(activity %in% 4:12, 0, 1))

any(is.na(training_data.t1$status)) ## making sure status column coded correctly


### Data for task 2
training_data.t2 <- training_data %>%
  mutate(status = ifelse(activity %in% 1:6, activity, 7))

any(is.na(training_data.t2$status)) ## making sure status column coded correctly
```

### Task 1

<b>Performing PCA</b>
```{r}
### getting correlation matrix
corr_matrix <- training_data %>% 
  dplyr::select(-c(1,2)) %>%
  cor()

### getting eigenvalues and eigenvectors
pca.data <- princomp(corr_matrix)
summary.data <- summary(pca.data)
pov <- summary.data$sdev^2/sum(summary.data$sdev^2) ## proportion of variance

pov[1:5]
```

<b>Regression Analysis</b>

Transform the data using the first three principal components
```{r}
### can make change here to which principal components to choose
transformed.t1.training <- as.matrix(dplyr::select(training_data.t1,-c(1,2,564))) %*% as.matrix(summary.data$loadings[,1:7]) %>%
  as.data.frame() %>%
  mutate(subject = training_data.t1$subject, activity = training_data.t1$activity, status = training_data.t1$status, .before = Comp.1)
```

Perform logistic regression
```{r}
task1.fit <- glm(data = transformed.t1.training, status ~ Comp.1 + Comp.2 + Comp.3 + Comp.4 + Comp.5 + Comp.6 + Comp.7, family = binomial(link = "logit"))

### get transformed test data
transformed.t1.testing <- as.matrix(dplyr::select(testing_data,-c(1))) %*% as.matrix(summary.data$loadings[,1:7]) %>%
  as.data.frame() %>%
  mutate(subject = testing_data$subject, .before = Comp.1)
```


Make predictions from model and checking preformance with the the training data set responses.
```{r}
task1.probs <- predict(task1.fit, transformed.t1.testing, type = "response")
task1.predict <- ifelse(task1.probs > 0.5, 1, 0)

check.probs <- predict(task1.fit, transformed.t1.training, type = "response")
check.predict <- ifelse(check.probs > 0.5, 1, 0)
mean(check.predict == transformed.t1.training$status)
```

Write .txt file
```{r}
# Writing predicted data
write.table(task1.predict, file = "binary_74892.txt", sep = "\t",
            row.names = FALSE, col.names = FALSE)
```

Summarizing performance with the training
```{r}
eval.tbl <- data.frame(check.predict) %>%
  mutate(`00` = ifelse(check.predict == 0 & check.predict == transformed.t1.training$status, TRUE, FALSE), 
         `11` = ifelse(check.predict == 1 & check.predict == transformed.t1.training$status, TRUE, FALSE),
         `01` = ifelse(check.predict == 0 & check.predict != transformed.t1.training$status, TRUE, FALSE),
         `10` = ifelse(check.predict == 1 & check.predict != transformed.t1.training$status, TRUE, FALSE))

colSums(eval.tbl[,2:5])/nrow(training_data)
```


### Task 2
<b>Regression Analysis</b>

```{r}
### can make change here to which principal components to choose
transformed.t1.training <- as.matrix(dplyr::select(training_data.t1,-c(1,2,564))) %*% as.matrix(summary.data$loadings[,1:11]) %>%
  as.data.frame() %>%
  mutate(subject = training_data.t1$subject, activity = training_data.t1$activity, status = training_data.t1$status, .before = Comp.1)

### get transformed test data
transformed.t1.testing <- as.matrix(dplyr::select(testing_data,-c(1))) %*% as.matrix(summary.data$loadings[,1:11]) %>%
  as.data.frame() %>%
  mutate(subject = testing_data$subject, .before = Comp.1)

### training data with multi-class response
transformed.t2.training <- transformed.t1.training %>%
  mutate(subject = training_data.t2$subject, activity = training_data.t2$activity, status = training_data.t2$status, .before = Comp.1)
  

task2.fit <- multinom(data = transformed.t2.training, status ~ Comp.1 + Comp.2 + Comp.3 + Comp.4 + Comp.5 + Comp.6 + Comp.7 + Comp.8 + Comp.9 + Comp.10 + Comp.11 )
```

Make predictions from model
```{r}
## seems better if I directlty have it predict type = class
task2.predict <- predict(task2.fit, transformed.t1.testing, type = "class")

check.predict <- predict(task2.fit, transformed.t2.training, type = "class")
```

Write .txt file
```{r}
# Writing predicted data
write.table(as.numeric(task2.predict), file = "multiclass_74894.txt", sep = "\t",
            row.names = FALSE, col.names = FALSE)
```


Creating figures displaying performance
```{r}
eval.tbl2 <- data.frame(check.predict,
                      transformed.t2.training$status, 
                      matches = (check.predict == transformed.t2.training$status)) 

ggplot(eval.tbl2, aes(x = factor(transformed.t2.training.status), y=(..count..)/sum(..count..), fill = matches)) +
  geom_bar(position = "dodge") +
  scale_y_continuous(labels = scales::percent, breaks = seq(0,16,by=0.015)) + 
  labs(title = "Correct/ Incorrect Predictions: Final Approach",
       y = "percentage(%)",
       x = "observed response",
       fill = "predicted response correct?") 
  
```